---
title: "HW3"
author: Yi Sun
output: github_document
---

Solution to HW3.

```{r setup, include = FALSE}
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
	fig.width = 6, 
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

Load the Instacart data.

```{r}
data("instacart")
```

The Instacart dataset contains `r nrows(instacart)` rows and `r ncol(instacart)` columns.

Each observation in this dataset is a product from an order submitted by a user.

In this dataset, there are variables about the user and order like user ID, order ID, order day, order hour, days since the last order, etc., and also variables about the product like product ID, product name, aisle, department, etc.

For example, one observation in the dataset tells us a product like Bulgarian Yogurt was bought in which order by which user (with IDs), on which day of the week, and which aisle and department this product belongs to, etc.

How many aisles are there, and which aisles are the most items ordered from?

```{r}
instacart %>% 
	count(aisle) %>% 
	arrange(desc(n))
```

There are 134 aisles, and the most items are ordered from the aisle "fresh vegetables".

Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize the plot.

```{r}
instacart %>% 
	count(aisle) %>% 
	filter(n > 10000) %>% 
	mutate(
		aisle = factor(aisle),
		aisle = fct_reorder(aisle, n)
	) %>% 
	ggplot(aes(x = aisle, y = n)) + 
	geom_point() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

The top 2 aisles are "fresh fruits" and "fresh vegetables".

Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in the table.

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(aisle, rank) %>% 
  knitr::kable()
```

The top 3 items in aisle "baking ingredients" are "Light Brown Sugar", "Pure Baking Soda", and "Cane Sugar".

The top 3 items in aisle "dog food care" are "Snack Sticks Chicken & Rice Recipe Dog Treats", "Organix Chicken & Brown Rice Recipe", and "Small Dog Biscuits".

The top 3 items in aisle "packaged vegetables fruits" are "Organic Baby Spinach", "Organic Raspberries", and "Organic Blueberries".

Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format the table as a 2 x 7 table.

```{r}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  )
```

The mean hour of the day at which Pink Lady Apples are ordered is generally earlier than that at which Coffee Ice Cream are ordered on each day of the week except on Fridays.

## Problem 2